### 2014-08-04 report/presentation.tex
Added final presentation.

\documentclass{beamer}
\usepackage{easyfig}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{CJK,CJKspace,CJKpunct}
\pdfmapline{=unisong@Unicode@ <ipam.ttf}
\easyfigdefault{placement={!htb}, max width=\textwidth, max height=.65\textheight, keepaspectratio}
\begin{document}
\begin{CJK}{UTF8}{zhsong}
\begin{frame}
\frametitle{JLyrics}
Larry Resnik and Zach Bian

CS 410: Introduction to Text Information Systems

August 2014
\end{frame}
\begin{frame}
\frametitle{Purpose}
%\framesubtitle{A bit more information about this}

\begin{center}
\begin{tabular}{l r}
Some Lyrics & Idiots who respond with "Dream" keep appearing \\
\hline
Title & Slow Motion \\
Artist & Pinochio P \\
Lyrics URL & http://piapro.jp/t/t30H \\
&\\
&\\
Some Lyrics & 「夢」と答えるボケを続けて \\
\hline
Title & すろぉもぉしょん \\
Artist & ピノキオP \\
Lyrics URL & http://piapro.jp/t/t30H \\
\end{tabular}
\end{center}
\end{frame}
\begin{frame}
\frametitle{Motivation}
\framesubtitle{Finding More Music}
\Figure[caption={Music by Pinochio P}]{pres-res/piapro-pinochio.jpg}
\end{frame}
\begin{frame}
\frametitle{Motivation}
\framesubtitle{Learning Japanese}
\Figure[caption={Rocket Player Showing Embedded Lyrics}]{res/rocket-player.png}
\end{frame}
\begin{frame}
\frametitle{Existing Work}
\framesubtitle{Anime Lyrics dot Com}
\Figure[caption={Anime Songs}]{pres-res/animelyrics.jpg}
\end{frame}
\begin{frame}
\frametitle{Existing Work}
\framesubtitle{Anime Lyrics dot Com}
\Figure[caption={Search for ``どのくらい知っている?''}]{pres-res/al_search0.jpg}
\end{frame}
\begin{frame}
\frametitle{Existing Work}
\framesubtitle{Anime Lyrics dot Com}
\Figure[caption={Search for ``どのくらい知っている?''}]{pres-res/al_search1.png}
\end{frame}
\begin{frame}
\frametitle{Existing Work}
\framesubtitle{Anime Lyrics dot Com}
\Figure[caption={Search for ``どのくらい知っている?''}]{pres-res/al_search2.png}
\end{frame}
\begin{frame}
\frametitle{Target Song Sources}
\framesubtitle{Niconico Douga}
\Figure[caption={Niconico Douga 24-Hour Ranking}, max height=.45\textheight]{pres-res/niconico0.jpg}

Most Representative of "Otaku" Interests.
\end{frame}
\begin{frame}
\frametitle{Target Song Sources}
\framesubtitle{Niconico Douga}
\Figure[caption={Niconico Douga 24-Hour Ranking}, max height=.45\textheight]{pres-res/niconico1.jpg}

Daily Rankings
\end{frame}
\begin{frame}
\frametitle{Target Song Sources}
\framesubtitle{Niconico Douga}
\Figure[caption={Niconico Douga 24-Hour Ranking}, max height=.45\textheight]{pres-res/niconico2.jpg}

Anime, Games, and Pictures Category
\end{frame}
\begin{frame}
\frametitle{Target Song Sources}
\framesubtitle{Niconico Douga}
\Figure[caption={Niconico Douga 24-Hour Ranking}, max height=.45\textheight]{pres-res/niconico3.jpg}

Sub-Categories: Anime, Games, Touhou, THE iDOLM@STER
\end{frame}
\begin{frame}
\frametitle{Target Song Sources}
\framesubtitle{Niconico Douga}
\Figure[caption={Niconico Douga 24-Hour Ranking}, max height=.45\textheight]{pres-res/niconico4.jpg}

Category: Entertainment and Music

Sub-Categories: Tried to Sing, Vocaloid, Niconico Indies
\end{frame}
\begin{frame}
\frametitle{Crawling}
\framesubtitle{Piapro}
\Figure[caption={Piapro Lyrics Section}]{pres-res/crawl_piapro0.png}
\end{frame}
\begin{frame}
\frametitle{Crawling}
\framesubtitle{Piapro}
\Figure[caption={Piapro Lyrics Section}]{pres-res/crawl_piapro1.png}
\end{frame}
\begin{frame}
\frametitle{Crawling}
\framesubtitle{Piapro}
\Figure[caption={Piapro Lyrics Section}]{pres-res/crawl_piapro2.png}
\end{frame}
\begin{frame}
\frametitle{Crawling}
\framesubtitle{Piapro}
\Figure[caption={Piapro Lyrics Section: Out of Bounds}]{pres-res/crawl_piapro3.png}
\end{frame}
\begin{frame}
\frametitle{Parse}
\framesubtitle{Piapro}
\Figure[caption={Lyrics to Pinocchio P's "Slow Motion"}]{pres-res/parse_piapro0.png}
\end{frame}
\begin{frame}
\frametitle{Crawling}
\framesubtitle{Anime Lyrics dot Com}
\Figure[caption={Top Level: Genres}]{pres-res/crawl_al0.png}
\end{frame}
\begin{frame}
\frametitle{Crawling}
\framesubtitle{Anime Lyrics dot Com}
\Figure[caption={Second Level: Albums}]{pres-res/crawl_al1.png}
\end{frame}
\begin{frame}
\frametitle{Crawling}
\framesubtitle{Anime Lyrics dot Com}
\Figure[caption={Third Level: Songs}]{pres-res/crawl_al2.png}
\end{frame}
\begin{frame}
\frametitle{Parsing}
\framesubtitle{Anime Lyrics dot Com}
\Figure[caption={Bottom Level: Individual Song}, max height=.45\textheight]{pres-res/parse_al0.png}

Translated and Transliterated Lyrics
\end{frame}
\begin{frame}
\frametitle{Parsing}
\framesubtitle{Anime Lyrics dot Com}
\Figure[caption={Bottom Level: Individual Song}, max height=.45\textheight]{pres-res/parse_al1.png}

Japanese Lyrics
\end{frame}
\begin{frame}
\frametitle{Indexing and Searching}

Goose: Base program using Lucene

Kuromoji: Japanese Analyzer
\end{frame}
\begin{frame}
\frametitle{Parse}
\framesubtitle{Piapro}
\Figure[caption={Lyrics to Pinocchio P's "Slow Motion"}]{pres-res/parse_piapro0.png}
\end{frame}
\begin{frame}
\frametitle{Indexing and Searching}
\framesubtitle{Piapro}
\Figure[caption={Source Code of Slow Motion Lyrics}]{pres-res/indexing_piapro0.png}
\end{frame}
\begin{frame}
\frametitle{Indexing and Searching}
\framesubtitle{Anime Lyrics dot Com}
\Figure[caption={Japanese Lyrics}]{pres-res/indexing0.png}
\end{frame}
\begin{frame}
\frametitle{Indexing and Searching}
\framesubtitle{Anime Lyrics dot Com}
\Figure[caption={Source Code}]{pres-res/indexing1.png}
\end{frame}
\begin{frame}
\frametitle{Indexing and Searching}
\framesubtitle{Anime Lyrics dot Com}
\Figure[caption={HTML Stripping}]{pres-res/indexing2.png}
\end{frame}
\begin{frame}
\frametitle{Indexing and Searching}
\framesubtitle{Anime Lyrics dot Com}

RE\_JAPANESE\_GLYPH = '[$ \backslash $p{Hiragana}$ \backslash $p{Katakana}$ \backslash $u4e00-$ \backslash $u9faf$ \backslash $W$ \backslash $u3005]'

\vspace*{1em}
spaced\_jp\_regex = re.compile(
    RE\_JAPANESE\_GLYPH + '$ \backslash $s+' +
    RE\_JAPANESE\_GLYPH, re.UNICODE)

\end{frame}
\begin{frame}
\frametitle{Indexing and Searching}
\framesubtitle{Anime Lyrics dot Com}
\Figure[caption={Proper Doc Format}]{pres-res/indexing3.png}
\end{frame}
\begin{frame}
\frametitle{Final Product}
\Figure[caption={JLyrics Running on Linux}]{res/410_post_Piapro.png}
\end{frame}
\begin{frame}
\frametitle{Final Product}
\Figure[caption={JLyrics Running on PuTTY}]{res/putty_search_success.png}
\end{frame}
\begin{frame}
\frametitle{Final Product}
\Figure[caption={JLyrics Running on Windows CMD}]{res/cmd_search_fail.png}
\end{frame}
\begin{frame}
\frametitle{Future Work}
\begin{itemize}
\item Windows support
\item Web page searching
\item Dataset expansion
\begin{itemize}
\item Anime Lyrics: translations and transliterations
\item Touhou Doujin CD no Kashi
\end{itemize}
\item Romaji $ \leftrightarrow $ Japanese text conversion \\ (synonyms such as toru, とる, and　取る)
\end{itemize}
\end{frame}
\end{CJK}
\end{document}